---
title: 'Baby steps: a simple illustration of the Metropolis algorithm'
author: "Steve Simon"
date: "2021-03-19"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Abstract

The Metropolis algorithm is a simple approach to generating random observations from a distribution where the density is known up to a proportionality constant. This algorithm is the easiest to understand and to implement of various methods used in Markov Chain Monte Carlo. In this paper, we show a simple example of the Metropolis algorithm to simulate the geometric distribution with $\pi$=1/2, and describe the resulting simulation as analogous to a baby learning to walk.

### Introduction

Consider a baby learning to walk. The baby is standing next to a couch, holding on for support. The baby decides to take one step away from the couch with probability 1/4 and stay next to the couch with probability 3/4. Once away from the security of the couch, the baby takes a step forward with probability 1/4, stays in place with probability 1/4, and steps backwards (towards the couch) with probability 1/2. How much time does the baby spend clutching the couch?

https://fivethirtyeight.com/features/will-the-baby-walk-away-will-the-troll-kill-the-dwarves/

You can solve this question easily with a computer simulation, but I derived the problem using the Metropolis algorithm to simulate a geometric distribution with p=1/2.

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(usmap)))

knitr::opts_chunk$set(echo=FALSE, fig.width=6, fig.height=3)
```

```{r}
sz <- 1.2
xc <- 20
yc <- 20
data.frame(x=20, y=20) %>%
  ggplot(aes(x, y)) +
  expand_limits(x=c(0, 100), y=c(0, 100)) +
  theme_void() +
  geom_text(x=xc, y=yc, label="0") +
  geom_text(x=xc+30, y=yc, label="1") +
  geom_text(x=xc, y=yc+30, label="p=3/4") +
  geom_text(x=xc+17.5, y=yc+30, label="p=1/4") +
  geom_curve(
    x=xc, y=yc+10, xend=xc-1, yend=yc+10, curvature=15, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+5, y=yc+10, xend=xc+30, yend=yc+10, curvature=-0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) -> sir_plot
sir_plot
```

*Figure 1. Baby walk

```{r}
sz <- 1.2
xc <- 20
yc <- 20
data.frame(x=20, y=20) %>%
  ggplot(aes(x, y)) +
  expand_limits(x=c(0, 100), y=c(0, 100)) +
  theme_void() +
  geom_text(x=xc, y=yc, label="k-1") +
  geom_text(x=xc+30, y=yc, label="k") +
  geom_text(x=xc+60, y=yc, label="k+1") +
  geom_text(x=xc+30, y=yc+30, label="p=1/4") +
  geom_text(x=xc+47.5, y=yc+30, label="p=1/4") +
  geom_text(x=xc+12.5, y=yc+30, label="p=1/2") +
  geom_curve(
    x=xc+30, y=yc+10, xend=xc+29, yend=yc+10, curvature=15, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+25, y=yc+10, xend=xc, yend=yc+10, curvature=0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+35, y=yc+10, xend=xc+60, yend=yc+10, curvature=-0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) -> sir_plot
sir_plot
```

*Figure 2. Baby walk

