---
title: 'Baby steps: a simple illustration of the Metropolis algorithm'
author: "Steve Simon"
date: "2021-03-19"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Abstract

The Metropolis algorithm is a simple approach to generating random observations from a distribution where the density is known up to a proportionality constant. This algorithm is the easiest to understand and to implement of various methods used in Markov Chain Monte Carlo. In this paper, we show a simple example of the Metropolis algorithm to simulate the geometric distribution with $\pi$=1/2, and describe the resulting simulation as analogous to a baby learning to walk.

### Introduction

Consider a baby learning to walk. The baby is standing next to a couch, holding on for support. The baby decides to take one step away from the couch with probability 1/4 and stay next to the couch with probability 3/4. Once away from the security of the couch, the baby takes a step forward with probability 1/4, stays in place with probability 1/4, and steps backwards (towards the couch) with probability 1/2. How much time does the baby spend clutching the couch?

https://fivethirtyeight.com/features/will-the-baby-walk-away-will-the-troll-kill-the-dwarves/

You can solve this question easily with a computer simulation, but I derived the problem using the Metropolis algorithm to simulate a geometric distribution with p=1/2.

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(usmap)))

knitr::opts_chunk$set(echo=FALSE, fig.width=6, fig.height=3)
```

```{r}
sz <- 1.2
xc <- 20
yc <- 20
data.frame(x=20, y=20) %>%
  ggplot(aes(x, y)) +
  expand_limits(x=c(0, 100), y=c(0, 100)) +
  theme_void() +
  geom_text(x=xc, y=yc, label="0") +
  geom_text(x=xc+30, y=yc, label="1") +
  geom_text(x=xc, y=yc+30, label="p=3/4") +
  geom_text(x=xc+17.5, y=yc+30, label="p=1/4") +
  geom_curve(
    x=xc, y=yc+10, xend=xc-1, yend=yc+10, curvature=15, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+5, y=yc+10, xend=xc+30, yend=yc+10, curvature=-0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) -> sir_plot
sir_plot
```

*Figure 1. Baby walk

```{r}
sz <- 1.2
xc <- 20
yc <- 20
data.frame(x=20, y=20) %>%
  ggplot(aes(x, y)) +
  expand_limits(x=c(0, 100), y=c(0, 100)) +
  theme_void() +
  geom_text(x=xc, y=yc, label="k-1") +
  geom_text(x=xc+30, y=yc, label="k") +
  geom_text(x=xc+60, y=yc, label="k+1") +
  geom_text(x=xc+30, y=yc+30, label="p=1/4") +
  geom_text(x=xc+47.5, y=yc+30, label="p=1/4") +
  geom_text(x=xc+12.5, y=yc+30, label="p=1/2") +
  geom_curve(
    x=xc+30, y=yc+10, xend=xc+29, yend=yc+10, curvature=15, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+25, y=yc+10, xend=xc, yend=yc+10, curvature=0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) +
  geom_curve(
    x=xc+35, y=yc+10, xend=xc+60, yend=yc+10, curvature=-0.6, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) -> sir_plot
sir_plot
```

*Figure 2. Baby walk

The Metropolis algorithm, first defined in Metropolis (1953) was originally developed to study the properties of substances composed of a finite number of interacting molecules. The method has proven useful for many physics problems such as simulated annealing (Kirkpatrick 1983)

### Discussion

John Kruschke motivates the Metropolis algorithm in Chapter 7 of his delightful book on Bayesian data analysis (Kruschke 2014). He describes an island hopping politician who wants to spend time on each island in proportion to its population. He can hop to the east or west, or stay on the current island, and uses the mechanics of the Metropolis algorithm to choose when to hop islands and when to stay put. The baby-walk example is similar in spirit, but provides a concrete distribution to evaluate.

Jim Albert and Jingchen Hu propose an arbitrary discrete distribution with five values and shows how students can simulate from this distribution using coin flips.

### Bibliography

S. Kirkpatrick; C. D. Gelatt; M. P. Vecchi
Science, New Series, Vol. 220, No. 4598. (May 13, 1983), pp. 671-680.

Kruschke, John (2014). Doing Bayesian Data Analysis. A Tutorial with R, JAGS, and Stan. 2nd edition. Academic Press.

Metropolis, Nicholas; Rosenbluth, Arianna W.; Rosenbluth, Marshall N.; Teller, Augusta H.; Teller, Edward (1953). "Equation of State Calculations by Fast Computing Machines". J. Chem. Phys. 21 (6): 1087. doi:10.1063/1.1699114.