---
title: "Hedging-hyperprior-plots"
author: "Steve Simon"
date: "June 19, 2017"
output: html_document
---

Just a quick note to myself:

Three dimensional plots are tricky. According to the Wikipedia entry on azimuth, "In mathematics, the azimuth angle of a point in cylindrical coordinates or spherical coordinates is the anticlockwise angle between the positive x-axis and the projection of the vector onto the xy-plane." The value of theta is defined as the azimuthal angle according to the help file of persp3D. So, assuming a colatitude (phi) of 0, the default of value of 0 for theta means that you are staring straight down the X-axis, and the value of 90 means that you are staring straight down the Y-axis.

```{r initial-steps}
source("prelims.R")
```

```{r draw-beta-binomial-nice}
png(file="plots/fig1.png", width=900, height=300)
layout(matrix(1:12, nrow=2, byrow=TRUE), 
       widths=rep(c(2, 8), 3)/30, heights=c(9, 1)/10)
par(mar=c(2.1, 2.1, 0.6, 0.6))
x <- (1:999)/1000
ct("dbeta(x, 4, 16)", 90)
pl2(dbeta(x, 4, 16))
ct("dbinom(18, 60, x)", 90)
pl2(dbinom(18, 60, x))
ct("dbeta(x, 22, 58)", 90)
pl2(dbeta(x, 22, 58))
par(mar=c(0.1, 2.1, 0.1, 0.6))
frame()
ct("Prior")
ct("X")
ct("Likelihood")
ct(expression(phantom(0) %prop% phantom(0)))
ct("Posterior")
dev.off()
```

```{r draw-beta-binomial-nasty}
png(file="plots/fig2.png", width=900, height=300)
layout(matrix(1:12, nrow=2, byrow=TRUE), 
       widths=rep(c(2, 8), 3)/30, heights=c(9, 1)/10)
par(mar=c(2.1, 2.1, 0.6, 0.6))
x <- (1:999)/1000
ct("dbeta(x, 4, 16)", 90)
pl2(dbeta(x, 4, 16))
ct("dbinom(54, 60, x)", 90)
pl2(dbinom(54, 60, x))
ct("dbeta(x, 58, 22)", 90)
pl2(dbeta(x, 58, 22))
par(mar=c(0.1, 2.1, 0.1, 0.6))
frame()
ct("Prior")
ct("X")
ct("Likelihood")
ct(expression(phantom(0) %prop% phantom(0)))
ct("Posterior")
dev.off()
```

```{r draw-hedging-surface}
png(file="plots/fig3.png", width=450, height=450)
th=120; ph=10
par(mar=rep(0.6, 4))
surf1 <- draw_prior()
dev.off()
```

```{r draw-tau-slices}
png(file="plots/fig5.png", width=450, height=450)
par(mar=rep(0.6, 4), mfrow=c(3, 3))
for (i in 8:0) {
  surf1 <- tau_slice((4*i+1)/34, g=35)
}
dev.off()
```

```{r draw-pi-slices}
png(file="plots/fig4.png", width=450, height=450)
par(mar=rep(0.6, 4), mfrow=c(3, 3))
for (i in 8:0) {
  surf1 <- pi_slice(i/8)
}
dev.off()
```

```{r draw-hedging-posterior-nasty}
x     <- 54
n     <- 60
g     <- 33
alpha <-  4
beta  <- 16
hp <- dens.uniform
png(file="plots/fig6.png", width=900, height=300)
par(mar=rep(0.6, 4))
layout(matrix(1:12, nrow=2, byrow=TRUE), 
       widths=rep(c(2, 8), 3)/30, heights=c(9, 1)/10)
frame()
s1 <- draw_prior(hp, g, alpha, beta)
frame()
s2 <- draw_likelihood(x, n, g)
frame()
s3 <- s1*s2
persp(z=s3, theta=th, phi=ph,
  xlab="tau", ylab="pi", zlab="Density")
frame()
ct("Prior")
ct("X")
ct("Likelihood")
ct(expression(phantom(0) %prop% phantom(0)))
ct("Posterior")
dev.off()
```

```{r draw-hedging-posterior-nice}
x     <- 18
n     <- 60
g     <- 33
alpha <-  4
beta  <- 16
hp <- dens.uniform
png(file="plots/fig7.png", width=900, height=300)
layout(matrix(1:12, nrow=2, byrow=TRUE), 
       widths=rep(c(2, 8), 3)/30, heights=c(9, 1)/10)
par(mar=rep(0.6, 4))
frame()
s1 <- draw_prior(hp, g, alpha, beta)
frame()
s2 <- draw_likelihood(x, n, g)
frame()
s3 <- s1*s2
persp(z=s3, theta=th, phi=ph,
  xlab="tau", ylab="pi", zlab="Density")
frame()
ct("Prior")
ct("X")
ct("Likelihood")
ct(expression(phantom(0) %prop% phantom(0)))
ct("Posterior")
dev.off()
```

```{r draw-hedge-trace}
png(file="plots/fig8.png", width=450, height=450)
hedge_trace <- plot_hedge_trace(101)
dev.off()
```

```{r draw-beta-hedging-surfaces}
par(mfrow=c(1,2))

hedge_trace1 <- plot_hedge_trace(101, hp=dens.beta12)
max(hedge_trace1$mn, na.rm=TRUE)

hedge_trace2 <- plot_hedge_trace(101, hp=dens.beta21)
max(hedge_trace2$mn, na.rm=TRUE)
hedge_trace2 %>% filter(mn>0.5) %>% slice(n())

hedge_trace3 <- plot_hedge_trace(101, hp=dens.beta19)
max(hedge_trace3$mn, na.rm=TRUE)

hedge_trace4 <- plot_hedge_trace(101, hp=dens.beta91)
max(hedge_trace4$mn, na.rm=TRUE)
hedge_trace4 %>% filter(mn>0.5) %>% slice(n())

par(mar=rep(0.6, 4), mfrow=c(1,2))
surf1 <- draw_prior(hp=dens.beta12,zmax=4.7)
surf1 <- draw_prior(hp=dens.beta21,zmax=9.4)
surf1 <- draw_prior(hp=dens.beta1T,zmax=11)
surf1 <- draw_prior(hp=dens.betaT1,zmax=47)
```


```{r save-everything}
save.image(file="hedge-plots.RData")
```